---
title: "Cultural Dynamics"
output:
  slidy_presentation:  
    css: font.css
header-includes: \usepackage{amsmath}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Wellcome to Slidy!

- Bullet 1
- Bullet 2
- Bullet 3

## A "simple" simulation of the game

```{r CoopCor, echo = TRUE}
CoopCor <- function(n_game,tmax,agent1,agent2,leader){
  
  #First, we create the main dataset we are going to export in the end
  #It is empty, with the right amount of rows and columns
  #We will add each run of the game to it
  game <-  as.data.frame(matrix(NA,n_game*tmax,10))
  colnames(game) <- c("P1_roll","P2_roll", "P1_cheat", "P2_cheat", "check","cheated", 
                      "payoff","th_payoff","ingame","n_game")
  
  #This overarching loop allows to play multiple games at once with 1 run of it being 1 game
  for (n in 1:n_game){
    
    #Then we define a still_in_game variable to identify teams who lost by GREED
    still_in_game <- 1
     
    #Then we can loop in time to play the game
      for (r in 1:tmax){
        
        #To Say where r should be in the dataframe (depend on the game)
        t = (n-1)*tmax + r
    
        
        ### P1 TURN
        #We draw between 1 and 6 using sample()
        game$P1_roll[t] <- sample(c(1:6),1)
        
        #If we don't have 6, we can decide to cheat, using agent1 STRATEGY (defined later)
        game$P1_cheat[t] <- ifelse(game$P1_roll[t]!= 6, agent1(game[1:t-1,]),0)
        game$P1_roll[t] <- ifelse(game$P1_cheat[t] == 1, 6, game$P1_roll[t])
        
        
        ### P2 TURN (following same process)
        game$P2_roll[t] <- sample(c(1:6),1)
        
        #if we are different from P1 we can decide to cheat, using agent2 STRATEGY
        game$P2_cheat[t] <- ifelse(game$P2_roll[t]!= game$P1_roll[t], agent2(game[1:t-1,]),0)
        game$P2_roll[t] <- ifelse(game$P2_cheat[t] == 1, game$P1_roll[t], game$P2_roll[t])
      
        #The following only works if they get the same diceroll
        x <- ifelse(game$P1_roll[t] == game$P2_roll[t],game$P1_roll[t],0)  
        game$th_payoff[t] <- x
        
        
        #### Leader watching over
        #Using a STRATEGY to decide to check or not (if agents succeeded only)
        game$check[t] <- ifelse(x != 0, leader(game[1:t-1,]),0)
        
        ### Payoff calculation
        #First we coarse-grain cheating(we don't need to know who did, just that it happened, and it is only relevant if they get a payoff)
        game$cheated[t] <- ifelse((game$P1_cheat[t] == 1 | game$P2_cheat[t] == 1) & x != 0, 1,0)
        #then we can run the equation (defined below)
        game$payoff[t] <- PayoffEq(x,game$check[t],game$cheated[t],still_in_game)
        
        #If a cheat happened, but no one checked, MENACING GAMBLE!!
        #a bernoulli with 5% chance to return TRUE, wich throws out of the game
        if(game$check[t] == 0 & game$cheated[t] == 1){
          if(rbernoulli(1,.05)){
            still_in_game <- 0}
          }
        
        #To finish, we report if we are still in game or if we lost this turn, and the game we are currently in
        game$ingame[t] <- still_in_game
        game$n_game[t] <- n
      } # this ends the loop for turns
    } # this ends the loop for games
  
  return(game)
}

#Equation to calculate payoff in synchronous payoff condition
#It takes in the dice roll value, wether there was a check or not, wether someone cheated or not, and wether the team is still in game or not
PayoffEq <- function(x,check,cheat,ingame){
  (x - check*x/2 - check*cheat*x/2)*ingame}
```

## Defining Strategies 1: Random

```{r, echo = TRUE}

BiasedChoice <- function(rate){
  # closure
  strategy <- function(game) {
    #We use a bernoulli to simulate a choice of 0 or 1 using rate as probability
    #As bernoulli returns TRUE and FALSE, we need to transform them in integers: 1 and 0
    return(as.integer(rbernoulli(1,rate)))
  }
  return(strategy)
}

```

## Time for some examples: What's the ideal?



## 


## Defining Strategies 2: Heuristics

Look at how much team is earning and make decision based on this
```{r, echo = TRUE}
#Has a memory of N + 1, looking at the mean payoff in a N + 1 window, compare with expectected mean payoff, if higher, check
N_Cumulative_Mean <- function(N){
  
  # closure
  strategy <- function(game) {
    #define the range to look for
    n_prev_rounds = nrow(game)
    begin_window = max(0, n_prev_rounds-N)
    
    #If we just started the game do it at random
    if (n_prev_rounds < 1) return(rbinom(1,1,0.5))
    
    # calculate how much team has earned in the last couple of rounds
    mean_payoff = mean(game$payoff[begin_window:n_prev_rounds])
    
    # this is what people should earn in general:
    expected_payoff = (1/6)*3.5 
    
    # Make a decision: if the mean payoff is higher than expected, check, if not don't
    strategic_choice = ifelse(mean_payoff > expected_payoff, 1, 0)
    
    # special case, if it's exactly on the cusp, do it random
    strategic_choice = ifelse(mean_payoff == expected_payoff, rbinom(1,1,.5), round(strategic_choice))
    
    return(strategic_choice)
  }
  return(strategy)
}

```

But wait, this is very lenient, as checked cheating will be counted as 0 (remember that teams don't get anything when their cheating is discovered). But they still cheated, so we may want to be a bit more grudgy.

```{r, echo = TRUE}
N_Th_Cumulative_Mean <- function(N){
  
  # closure
  strategy <- function(game) {
    #define the range to look for
    n_prev_rounds = nrow(game)
    begin_window = max(0, n_prev_rounds-N)
    
    #If we just started the game do it at random
    if (n_prev_rounds < 1) return(rbinom(1,1,0.5))
    
    mean_payoff = mean(game$th_payoff[begin_window:n_prev_rounds]) # we're taking the theoretic payoff, rather than the actual payoff
    expected_payoff = (1/6)*3.5 
    strategic_choice = ifelse(mean_payoff > expected_payoff, 1, 0)
    strategic_choice = ifelse(mean_payoff == expected_payoff, rbinom(1,1,.5), round(strategic_choice))
    
    return(strategic_choice)
  }
  return(strategy)
}

```


## Defining Strategies 3: Refining Heuristics

So far, we've been very strict: either you're above the limit or not. But this is probably not a good idea. After all, world is inherently uncertain, and scoring above the expected mean could also be due to chance. Maybe we should be a bit more lenient.... If only we had a way to test whether an event is due to chance or not, given a theoretical distribution...

```{r, echo = TRUE}
N_Cumulative_Mean_ttest <- function(N){
  
  # closure
  strategy <- function(game) {
    #define the range to look for
    n_prev_rounds = nrow(game)
    begin_window = max(0, n_prev_rounds-N)
    prev = game$payoff[begin_window:n_prev_rounds]
   
    ##A t-test cannot handle less than 2 datapoint to compare, and cannot handle no variance
    #If we just started the game do it at random until we have a history of at least size 2
    if (n_prev_rounds < 2) return(rbinom(1,1,0.5))
    
    #If all values in the window are the same:
    #it is only 0s, in which case mean = 0, so don't check
    #it is only 1 or 2 or..., in which case mean >> 0.58, so check
    if (length(unique(prev)) == 1) return(as.integer(prev[1] > 0))

    expected_payoff = (1/6)*3.5 
    # print(expected_payoff)
    p_check = 1 - t.test(prev,mu = expected_payoff, alternative = "greater")$p.value
    strategic_choice = as.integer(rbernoulli(1,p_check))
    
    return(strategic_choice)
  }
  return(strategy)
}
```

Again, we can be a bit more grudgy...
```{r}
N_Th_Cumulative_Mean_ttest <- function(N){
  
  # closure
  strategy <- function(game) {
    #define the range to look for
    n_prev_rounds = nrow(game)
    begin_window = max(0, n_prev_rounds-N)
    prev = game$th_payoff[begin_window:n_prev_rounds]

    ##A t-test cannot handle less than 2 datapoint to compare, and cannot handle no variance
    #If we just started the game do it at random until we have a history of at least size 2
    if (n_prev_rounds < 2) return(rbinom(1,1,0.5))
    
    #If all values in the window are the same:
    #it is only 0s, in which case mean = 0, so don't check, it's probably fine
    #it is only 1 or 2 or..., in which case mean >> 0.58, so check, it's suspicious
    if (length(unique(prev)) == 1) return(as.integer(prev[1] > 0))

    expected_payoff = (1/6)*3.5 
    p_check = 1 - t.test(prev,mu = expected_payoff, alternative = "greater")$p.value
    strategic_choice = as.integer(rbernoulli(1,p_check))
    
    return(strategic_choice)
  }
  return(strategy)
}
```


## Time for some visualisations
